# Data transformation
Since we will only extract some information from the two auxiliary data sources to help analyze our main dataset, we mainly discuss the data transformation of the main dog bites dataset.

## The column “DateOfBite”

In our main dog bites dataset, the data type of the column named “DateOfBite” is “Date & Time”, e.g., “January 02, 2015”. In our analysis, we aim to discuss the passage of time in months, thus we extract the “year” and “month” from this column to add two new columns named “MonthOfBite” and “YearOfBite”. We show the original “DataOfBite” and the two new columns for some rows as follows.

```{r, message=FALSE}
library(pacman)
p_load(tidyverse)
p_load(lubridate)
p_load(ggExtra)


lct <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_TIME", "C")

DogBiteDF <- read_csv("DOHMH_Dog_Bite_Data.csv", na = c("", "NA","?"))
Temperature <- read_csv("tempbymonth.csv", na = c("", "NA"))
DogRegDF <- read_csv("NYC_Dog_Licensing_Dataset.csv")
DogBiteDF <- DogBiteDF %>% 
  mutate(MonthOfBite = month(strptime(pull(DogBiteDF, DateOfBite), 
                                      format="%B %d %Y"))) %>% 
  mutate(YearOfBite = year(strptime(pull(DogBiteDF, DateOfBite), 
                                      format="%B %d %Y")))
DogBiteDF[, c("DateOfBite", "MonthOfBite", "YearOfBite")]
```

## The column “Breed”

Many rows contain an empty “Breed”, we transform them as “UNKNOWN”. We show the “Breed” column and it successfully shows “UNKNOWN” in some rows.

```{r}
DogBiteDF <- DogBiteDF %>% 
  mutate(Breed = replace_na(Breed, 'UNKNOWN'))
DogBiteDF[, c("Breed")]
```

## Temperature data from Average Temperature in NYC Data

We use our Average Temperature in NYC Data to obtain the monthly average temperature for the corresponding month of each dog bite. We use “MonthOfBite” and “YearOfBite” to match our rows to the corresponding temperature, and add the corresponding temperature as a new column named “Avg_temp” to the main dog bites dataset. We show the “Avg_temp”, “MonthOfBite” and “YearOfBite” columns as follows.

```{r}
Temperature <- Temperature %>% 
  pivot_longer(!Year, names_to = "Month", 
               names_transform = list(Month = as.integer),
               values_to = "Avg_temp",
               values_drop_na = TRUE)

DogBiteDF <- DogBiteDF %>% left_join(., Temperature, 
                                     by=c('MonthOfBite'='Month', 'YearOfBite'='Year'), suffix = c("_x", "_y"))
DogBiteDF[, c("Avg_temp","MonthOfBite","YearOfBite")]
```

## Bourough data in NYC Dog Licensing Dataset

In our auxiliary dataset (NYC Dog Licensing Dataset), the values of “Borough” is missing for all of the rows, thus we use the column “ZipCode” to classify the records to the corresponding boroughs. We queried the range of ZipCodes in NYC corresponding to each borough, and marked each row with the corresponding “Borough” according to its “ZipCode”. We show the “ZipCode” and “Borough” columns as follows.

```{r}
DogRegDFtt = DogRegDF
DogRegDFtt $ZipCode = as.numeric(DogRegDFtt $ZipCode)
DogRegDFtt$Borough[DogRegDFtt$ZipCode>=10001 & DogRegDFtt$ZipCode <= 10282] = "Manhattan"
DogRegDFtt$Borough[DogRegDFtt$ZipCode>=10451 & DogRegDFtt$ZipCode <= 10475] = "Bronx"
DogRegDFtt$Borough[DogRegDFtt$ZipCode>=11201 & DogRegDFtt$ZipCode <= 11256] = "Brooklyn"
DogRegDFtt$Borough[(DogRegDFtt$ZipCode>=11004 & DogRegDFtt$ZipCode <= 11109) | (DogRegDFtt$ZipCode>=11351 & DogRegDFtt$ZipCode <= 11697)] = "Queens"
DogRegDFtt$Borough[DogRegDFtt$ZipCode>=10301 & DogRegDFtt$ZipCode <= 10314] = "Staten Island"
LicT = DogRegDFtt[c("Borough")]
LicT$group = "Licensed"
BiteT = DogBiteDF[c("Borough")]
BiteT$group = "Biting"
data_for_draw8 = rbind(LicT, BiteT)
DogRegDFtt[, c("ZipCode","Borough")]
```
